Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-08-18 Ben_Manuscripts/structure_paper/Draft.tex
comments.

% BJC6: removed to cut down on amount of material. This one is the biggest stretch in my opinion.
%MRS12: I think it is useful to put in just to show we've been thinking about them.  Can omit the figure in the main text.
 One way toproduce R-double if our initial configuration contains alternating parallel and
anti-parallel carboxylate groups relative to the plane of the monomer's phenyl ring.
%  (Figure ~\ref{fig:rotated_carboxylate_rzplot_norestraints}). 
It is difficult to physically justify this system. Systems built this way are only stable if position restraints are placed on all head group heavy atoms. Carboxylate groups quickly revert to the parallel position as restraints are released. There is an appreciable energy barrier that prevents rotation of carboxylate groups attached to phenyl rings since the group extends the system's $\pi$ conjugation (citation) (See Figure S-\ref{S-fig:carboxylate_dihedral_rb}), which is significant even considering possible overestimates of the barrier height in the classial force field. 
%  
%  %BJC5: add citation for extended pi conjugation
%  There are instances where carboxylate groups in other systems rotate out of plane. Bushey et
%  al. showed that bulky carbonyl-containing substituents of stacked arene molecules tended to
%  tilt 45$\degree$ in order to relieve steric strain, thus allowing $\pi$-$\pi$ stacking between 
%  molecules, and to hydrogen bond with neighboring molecules. Lorenzo and Gra\~{n}a found 3
%  stable dimers of gallic acid (from which the monomer, Na-GA3C11 is derived) with adenine. In
%  all cases, the carboxylate group remained planar, even with evidence of hydrogen bonding
%  between the carboxyl group and nitrogen atoms of adenine. The monomers which we are studying
%  have no opportunities to hydrogen bond and are confined so that any rotation about the phenyl--
%  carboxylate bond would be sterically hindered.

@@

  We can also produce R-double if layers are not uniformly, but instead form pairs that stack less 
  than 3.7~\AA~apart, with center
  of masses are spaced 7.4 \AA~ from neighboring pairs of monomers (Figure
  ~\ref{fig:staggered_rzplot_norestraints}). 
  %MRS12: not sure this is necessary.
  %To our knowledge, there have been no
  %studies that specifically address the possibility of a configuration like this.
  Our force field causes our system to tend towards uniformly spaced layers.
  Simulations of unevenly spaced systems are only stable if position restraints
  are applied to heavy atoms of the phenyl rings.  Additionally, there is little evidence from
  QM studies of stacked $\pi-\pi$ systems that such uneven stacking could be energetically stable.
  incorrect.~\ref{tauer_beyond_2005}

@@

  \begin{figure}[!htb]
  \begin{subfigure}{0.45\textwidth}
  \centering
  % BJC6 : below is the simulated diffraction pattern of the full system. I removed it since showing both
  % makes both patterns small. You can see the full pattern in a separate figure.
%  \begin{subfigure}{0.45\textwidth}
%  \includegraphics[width=\textwidth]{rdouble_offset_solvated_rzplot.png}  
%  \end{subfigure}
  \begin{subfigure}{\textwidth}
  \includegraphics[width=\textwidth]{nophenyls_rzplot.png}
  \end{subfigure}
  \caption{}\label{fig:rdouble_nophenyls}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{hbonds_tails.png}
  \caption{}\label{fig:hbond_visualization}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{pore_hbonds.png}
  \caption{}\label{fig:pore_hbonds}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{pore_hbonds_ft.png}
  \caption{}\label{fig:pore_hbonds_ft}
  \end{subfigure}
  \caption{(a) The structure of the head groups is responsible for the appearance of
  R-double. When we remove head groups from the trajectory, the simulated 
  diffraction pattern no longer shows R-double. (b) Monomer head groups above or below
  each other that hydrogen bond with a shared water molecule are drawn closer
  together in the $z$-direction. Blue monomers were stacked above orange monomers in the
  initial configuration. (c) We group the $z$-coordinate of the center of mass of monomer
  head groups into 20 bins. Peaks in the distributions indicate a shared hydrogen bonded
  water molecule between a monomer head group and one in the bin vertically above it. 
  The distance between the peaks roughly corresponds to the distance between pairs of 
  monomers. The distance between peaks in pore 1 (indicated by arrows) is equal to two 
  times the $\pi$-stacking distance. In this scenario, monomers are no longer evenly spaced
  in the $z$-direction since they are pulled closer together by hydrogen bonds. (d) We 
  performed discrete Fourier transforms on each distribution in (c) In Pore 1, there 
  also periodicity every 8.3 $\AA$, twice the $\pi$-stacking distance, as indicated by the
  intensity at 0.76 $\AA^{-1}$). In this case, the center of mass of two adjacent pairs 
  is separated by twice the average monomer stacking distance, while the four monomers 
  involved are unequally spaced. These conditions may give rise to R-double.}\label{fig:hbonds}
  \end{figure}

>>>>

% BJC6: removed to cut down on amount of material. This one is the biggest stretch in my opinion.
%MRS12: I think it is useful to put in just to show we've been thinking about them.  Can omit the figure in the main text.
 One way to produce R-double if our initial configuration contains alternating parallel and
anti-parallel carboxylate groups relative to the plane of the monomer's phenyl ring.
%  (Figure ~\ref{fig:rotated_carboxylate_rzplot_norestraints}). 
It is difficult to physically justify this system. Systems built this way are only stable if position restraints are placed on all head group heavy atoms. Carboxylate groups quickly revert to the parallel position as restraints are released. There is an appreciable energy barrier that prevents rotation of carboxylate groups attached to phenyl rings since the group extends the system's $\pi$ conjugation (citation) (See Figure S-\ref{S-fig:carboxylate_dihedral_rb}), which is significant even considering possible overestimates of the barrier height in the classial force field. 
%  
%  %BJC5: add citation for extended pi conjugation
%  There are instances where carboxylate groups in other systems rotate out of plane. Bushey et
%  al. showed that bulky carbonyl-containing substituents of stacked arene molecules tended to
%  tilt 45$\degree$ in order to relieve steric strain, thus allowing $\pi$-$\pi$ stacking between 
%  molecules, and to hydrogen bond with neighboring molecules. Lorenzo and Gra\~{n}a found 3
%  stable dimers of gallic acid (from which the monomer, Na-GA3C11 is derived) with adenine. In
%  all cases, the carboxylate group remained planar, even with evidence of hydrogen bonding
%  between the carboxyl group and nitrogen atoms of adenine. The monomers which we are studying
%  have no opportunities to hydrogen bond and are confined so that any rotation about the phenyl--
%  carboxylate bond would be sterically hindered.

@@

  We can also produce R-double if layers are not uniformly, but instead form pairs that stack less 
  than 3.7~\AA~apart, with center
  of masses are spaced 7.4 \AA~ from neighboring pairs of monomers (Figure
  ~\ref{fig:staggered_rzplot_norestraints}). 
  %MRS12: not sure this is necessary.
  %To our knowledge, there have been no
  %studies that specifically address the possibility of a configuration like this.
  Our force field causes our system to tend towards uniformly spaced layers.
  Simulations of unevenly spaced systems are only stable if position restraints
  are applied to heavy atoms of the phenyl rings.  Additionally, there is little evidence from
  QM studies of stacked $\pi-\pi$ systems that such uneven stacking could be energetically stable.~\ref{tauer_beyond_2005}

@@

  \begin{figure}[!htb]
  \begin{subfigure}{0.45\textwidth}
  \centering
  % BJC6 : below is the simulated diffraction pattern of the full system. I removed it since showing both
  % makes both patterns small. You can see the full pattern in a separate figure.
%  \begin{subfigure}{0.45\textwidth}
%  \includegraphics[width=\textwidth]{rdouble_offset_solvated_rzplot.png}  
%  \end{subfigure}
  \begin{subfigure}{\textwidth}
  \includegraphics[width=\textwidth]{nophenyls_rzplot.png}
  \end{subfigure}
  \caption{}\label{fig:rdouble_nophenyls}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \centering
  \includegraphics[width=\textwidth]{hbonds_tails.png}
  \caption{}\label{fig:hbond_visualization}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{pore_hbonds.png}
  \caption{}\label{fig:pore_hbonds}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{pore_hbonds_ft.png}
  \caption{}\label{fig:pore_hbonds_ft}
  \end{subfigure}
  \caption{(a) The structure of the head groups is responsible for the appearance of
  R-double. When we remove head groups from the trajectory, the simulated 
  diffraction pattern no longer shows R-double. (b) Monomer head groups above or below
  each other that hydrogen bond with a shared water molecule are drawn closer
  together in the $z$-direction. Blue monomers were stacked above orange monomers in the
  initial configuration. 
  %MRS12: I think subfigure C can be moved to supporting, and we just desribe the alternating structure in the text.  I think %it's likely to be too confusing for many readers as a figure, and takes up too much space.
(c) We group the $z$-coordinate of the center of mass of monomer
  head groups into 20 bins. Peaks in the distributions indicate a shared hydrogen bonded
  water molecule between a monomer head group and one in the bin vertically above it. 
  The distance between the peaks roughly corresponds to the distance between pairs of 
  monomers. The distance between peaks in pore 1 (indicated by arrows) is equal to two 
  times the $\pi$-stacking distance. In this scenario, monomers are no longer evenly spaced
  in the $z$-direction since they are pulled closer together by hydrogen bonds. (d) We 
  performed discrete Fourier transforms on each distribution in (c) In Pore 1, there 
  also periodicity every 8.3 $\AA$, twice the $\pi$-stacking distance, as indicated by the
  intensity at 0.76 $\AA^{-1}$). In this case, the center of mass of two adjacent pairs 
  is separated by twice the average monomer stacking distance, while the four monomers 
  involved are unequally spaced. These conditions seem likely to give rise to R-double.}\label{fig:hbonds}
  \end{figure}
